\documentclass{report}

\input{preamble}
\input{macros}
\input{letterfonts}
\newcommand{\contra}{
$\rightarrow\!\leftarrow$
}

\title{Ultimate Problem Set}
\author{Jack Krebsbach }

\date{Dec 11th}

\begin{document}
\maketitle



\qs{}{

Suppose that $x>-1$ and that $x \neq 0$. Prove that
$$
(1+x)^n>1+n x
$$
for each integer $n>1$. This result is know as Bernoulli's inequality.
}
\qs{}{

Show that $e$ is irrational by supposing that $e=\frac{m}{n}$ and deriving a contradiction. Use the fact that $e=\sum_{j=0}^{\infty} \frac{1}{j !}$. Let $s_k=\sum_{j=0}^k \frac{1}{j !}$.
}

(a) Prove that
$$
e-s_k<\frac{1}{(k+1) !}\left\{1+\frac{1}{k+1}+\left(\frac{1}{k+1}\right)^2+\cdots\right\} .
$$
(b) Prove that $e-s_k<\frac{1}{k(k+)}$ for all $k \in \mathbb{N}$.
(c) If $e=\frac{m}{n}$, prove that $n ! e$ and $n ! s_n$ are integers.
(d) If $e=\frac{m}{n}$, prove that $n !\left(e-s_n\right)$ is an integer between 0 and 1 , which is absurd.



\qs{}{
Let $f$ be a function defined on all of $\mathbb{R}$, and assume there is a constant $c$ such that $0<c<1$ and
$$
|f(x)-f(y)| \leq c|x-y|
$$
for all $x, y \in \mathbb{R}$.
}

(a) Show that $f$ is continuous.
(b) Pick some $y_1 \in \mathbb{R}$ and construct the sequence
$$
\left(y_1, f\left(y_1\right), f\left(f\left(y_1\right)\right), \ldots\right) .
$$

In general, if $y_{n+1}=f\left(y_n\right)$, show that the resulting sequence $\left(y_n\right)$ is a Cauchy sequence. Hence we may let $y=\lim _{n \rightarrow \infty} y_n$.
(c) Prove that $y$ is a fixed point of $f$ (i.e. $f(y)=y$ ) and that it is unique in this regard.
(d) Finally, prove that if $x$ is any arbitrary point in $\mathbb{R}$, then the sequence $(x, f(x), f(f(x)), \ldots)$ converges to $y$ (as defined in (b)).


\qs{}{
Let $\left\{r_n\right\}$ be a listing of all the rational numbers. Define a function $f$ by $f(x)=0$ if $x$ is irrational and $f\left(r_n\right)=1 / n$ for all $n$. Show that $f$ is continuous everywhere except for the set of rational numbers.
}

\qs{}{

Using the $\delta-\epsilon$ definition of a limit, show
$$
\lim _{x \rightarrow 1} \frac{x^3-1}{x-1}=3 .
$$
}





\end{document}
